---
title: "Ethnic Differences in Phonetic Production"
author: "Ryan C McMurry"
format: pptx
editor: visual
---

## Introduction

Are their phonetic production differences along ethnic lines?

-   If yes, do these differences show in vowel production?

-   How do these differences pattern along current vowel shifts?

## Caveats

What I do not have time to discuss

-   Definition of Ethnicity
    -   Ethnicity is a very fluid human construction
    -   No two people will have the exact same ethnic expression
-   Despite this there are general patterns that can be found
    -   Social Identity Theory and social significance and meaning (Hogg, 2016)
    -   This meaning is conveyed by shared conventions (Ochs, 1993)
    -   Conventions are shared by concrete sounds (Eckert & Labov, 2017)
-   Phonetic NOT Phonemic
    -   I am looking at this from a phonetic lens

## Previous Work

Vowel differences along ethnic lines

-   Variation of Northern City Shift (Gordon, 2000)

    -   Raising of /æ/ was not seen in African and Mexican American population

    -   Raising of /ɛ/ was not seen in the African and Mexican American population

    -   White Americans appeared to lead the shift

## Previous Work (cont.)

Vowel differences along ethnic lines

-   Variation in San Francisco (Cardoso et al. 2016)

    -   Differences in nasality in /æ/

    -   Chinese Americans showed less nasality in words like BAN

    -   White Americans generally produced more nasality

## Current Study

Will I find vowel differences from the students at UC Davis?

-   22 participants (18-20 yo)

    -   12 East Asian (ethnically descendant of Japanese, Chinese, and Philippine peoples)

    -   10 White (ethnically descendant of European peoples)

-   From California

-   HvD, BvD, BvT, Pen/Pin, two diphthongs, pull/pool

## Low Back Merger (LBM)

Merger throughout North America of merging of /ɔ/ and /ɑ/ vowels (Doernberger & Cerny, 2008)

-   COT and CAUGHT

California

-   fronting of /u/ and the lowering of /I/ and /ɛ/

-   raising of /æ/ in some contexts

## Method

-   Participants were recorded using Praat (44100 hertz)

-   Annotated the wav files and aligned using DARLA (Dartmouth Linguistic Automation)

-   Extracted the vowels using FastTrack (Barreda, 2021)

-   Normalized the vowels using log-mean normalization (Barreda & Nearey, 2018)

## Method (Cont.)

Linear Discriminant Analysis & Signal Detection Theory Analysis

-   LDA: Can a model to be trained to the formants of each ethnicity?

    -   Predictions could suggest where in the vowel space each ethnicity produces the vowel

-   SDT: How accurate and reliable is that model?

    -   What is the hit rate and false alarm rate?

## LDA Predictions

```{r, include=FALSE}
library(tidyverse)
library(ggplot2)
library(MASS)
library(dplyr)
library(ggrepel)
lda_data = readxl::read_excel("C:\\Users\\verde\\OneDrive\\Desktop\\Prelim\\Prelim\\data.xlsx")
lda_data=lda_data[lda_data$vowel!="i",]
lda_model <- lda(vowel_ethnicity ~ f11 + f21 + f31 + f15 + f25 + f35, 
                 data = lda_data)
predictions <- predict(lda_model)
df_actual <- lda_data %>%
  group_by(vowel_ethnicity) %>%
  summarise(
    f1_avg = mean(rowMeans(across(c(f11, f15)), na.rm = TRUE)),
    f2_avg = mean(rowMeans(across(c(f21, f25)), na.rm = TRUE)),
    type = "Actual"
  )

# Compute predicted vowel-ethnicity averages
df_predicted <- lda_data %>%
  group_by(predicted_ethnicity, vowel) %>%  # Need vowel for clarity in grouping
  summarise(
    f1_avg = mean(rowMeans(across(c(f11, f15)), na.rm = TRUE)),
    f2_avg = mean(rowMeans(across(c(f21, f25)), na.rm = TRUE)),
    type = "Predicted"
  ) 

# Combine both sets into one dataset
df_plot01 <- bind_rows(df_actual, df_predicted)
df_plot_clean <- df_plot01 %>%
  filter(!is.na(vowel), !is.na(predicted_ethnicity))

```

lda_model \<- lda(vowel_ethnicity \~ f11-f15+f21-f25+f31-f35)

Divide the vowel space into 3

1.  Front Vowels

    Differ in terms of "height"

    The most dramatic /eI/ and /I/ while /ɛ/ is less so

2.  Mid Vowels

    Difficult to generalize a pattern

    Perhaps /ɝ/ is different

    Vowels may have great overlap in production

3.  Back Vowels

    Differ in terms of "frontness"

    /u/ and /ʊ/ strikingly different

```{r, echo=FALSE}
ggplot(df_plot_clean, aes(x = f2_avg, y = f1_avg, color = vowel, shape = predicted_ethnicity)) +
  geom_point(size = 4, alpha = 0.8) +  
  geom_text_repel(aes(label = vowel), size = 5, fontface = "bold") +  
  scale_y_reverse() +  
  scale_x_reverse() +  
  theme_classic() +
  labs(
    title = "LDA Prediction of Vowels by Ethnicity",
    x = "F2 (Hz)",
    y = "F1 (Hz)",
    color = "Vowel",
    shape = "Predicted Ethnicity"
  ) +
  theme(
    legend.position = "bottom",
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12)
  )
```

## LBM Context

Low back vowels are merged (how surprising)

Differences in the fronting /u/ and /ʊ/

-   The White population follows this trend more

Differences in /ɪ/, and /ɛ/ lowering

-   "Higher" in the EA population

-   The W population follows this trend more

```{r, echo=FALSE}
ggplot(df_plot_clean, aes(x = f2_avg, y = f1_avg, color = vowel, shape = predicted_ethnicity)) +
  geom_point(size = 4, alpha = 0.8) +  
  geom_text_repel(aes(label = vowel), size = 5, fontface = "bold") +  
  scale_y_reverse() +  
  scale_x_reverse() +  
  theme_classic() +
  labs(
    title = "LDA Prediction of Vowels by Ethnicity",
    x = "F2 (Hz)",
    y = "F1 (Hz)",
    color = "Vowel",
    shape = "Predicted Ethnicity"
  ) +
  theme(
    legend.position = "bottom",
    axis.title = element_text(size = 14, face = "bold"),
    axis.text = element_text(size = 12)
  )
```

## SDT Analysis

Front Vowels

-   /eɪ/ in the White population low hit rate /eɪ/ high rate

-   /ɛ/ high hit rate in both populations

-   /ɪ/ high hit rate in both populations

Back Vowels

-   /ʊ/ fifty-fifty split in the W population high in the EA population

-   /u/ high hit rate in both populations

Can we trust these results?

```{r, include=FALSE}
df_sdt <- lda_data %>%
  mutate(
    hit = ifelse(ethnicity == predicted_ethnicity, 1, 0),
    false_alarm = ifelse(ethnicity != predicted_ethnicity, 1, 0)
  ) %>%
  group_by(vowel_ethnicity) %>%
  summarise(
    hit_rate = mean(hit),
    false_alarm_rate = mean(false_alarm),
    .groups = "drop"
  )
df_sdt <- df_sdt %>%
  mutate(
    d_prime = qnorm(hit_rate) - qnorm(false_alarm_rate)
  )
df_sdt <- df_sdt %>%
  mutate(
    response_bias = -0.5 * (qnorm(hit_rate) + qnorm(false_alarm_rate))
  )
library(pROC)
data = data.frame(
  vowel_ethnicity = c("eɪ_EA", "eɪ_W", "u_EA", "u_W", "ɛ_EA", "ɛ_W","ɪ_EA", "ɪ_W", "ʊ_EA", "ʊ_W"),
hit_rate = c(0.7778, 0.4667,0.6458, 0.6125,0.7845, 0.5305,0.7924, 0.5335, 0.8261, 0.5 ),
false_alarm_rate = c(0.2222, 0.5333,0.3542, 0.3875,0.2155, 0.4695,0.2076, 0.4665, 0.1739, 0.5 ))
```

```{r, echo=FALSE}
data <- data[order(data$false_alarm_rate), ]
ggplot(data, aes(x = false_alarm_rate, y = hit_rate, color = vowel_ethnicity)) +
  geom_point(size=5) +
  labs(title = "ROC Curve for Vowel Classification",
       x = "False Alarm Rate",
       y = "Hit Rate") +
  theme_minimal()
```

## SDT Analysis (Cont.)

Yes, and

Overall d' Values (sensitivity) relatively moderate \~1.5

-   Suggest the mode is able to properly discriminate between the two ethnicities

-   However, some overlap does exist

Lower d' Values in the W population

-   d' for the /ɪ/ vowel in the White population only .16

-   d' for the /u/ vowel in both populations only .75 and .57 (EA & W)

Perhaps there is more overlap within groups then between?

```{r, echo=FALSE}
data <- data[order(data$false_alarm_rate), ]
ggplot(data, aes(x = false_alarm_rate, y = hit_rate, color = vowel_ethnicity)) +
  geom_point(size=5) +
  labs(title = "ROC Curve for Vowel Classification",
       x = "False Alarm Rate",
       y = "Hit Rate") +
  theme_minimal()
```

## Thoughts

Production differences appear to exist

Front vowels

-   Predicted to be higher in the EA population than the W

Back vowels

-   Predicted to be more fronted in the W population than the EA

Intra-group Variation

-   d' values overall lower in the EA population

-   Possible evidence of greater variation in the W population

## Conclusion

MORE DATA!

-   In order to increase accuracy and raise d' values

## Thank you
